Loading and Preprocessing data
========================================================
```{r}
activity<-read.csv("E:/Coursera/RWorkDir/activity.csv")
library(plyr)
activity1<-ddply(activity,.(date),summarise,steps=sum(as.numeric(steps)))
library(lubridate)
activity1$date<-strptime(activity1$date,"%Y-%m-%d")
```

## Total number of steps taken per day.
### Plot.

```{r fig.width=11, fig.height=6}
plot(activity1$date,activity1$steps,type="h",lwd=11,col="dark red",lty=1,
   lend=1,main="No.of Steps taken per day.",axes=F,xlab=NA,ylab=NA)
axis.POSIXct(1,format="%d %b",at=activity1$date,labels=TRUE,las=2,tck=-.01)
axis(side=2,las=2,tck=-.01)
mtext(side = 1, "Days", line = 4)
mtext(side = 2, "Steps", line = 3.2)
```

### Mean & Median .
```{r}
Stepsmean<-mean(activity1$steps,na.rm=TRUE)
```

The mean of steps taken per day is `r Stepsmean`(10766.19)

```{r}
Stepsmedian<-median(activity1$steps,na.rm=TRUE)
```

The mean of steps taken per day is `r Stepsmedian`(10765).

## The average daily activity pattern.

```{r}
activity2<-na.omit(activity)
library(plyr)
activity3<-ddply(activity2,.(interval),summarise,steps            =mean(as.numeric(steps)))
```

### Plot.

```{r}
plot(activity3$interval,activity3$steps,type="l",lwd="1",
     col="dark red",xlab="Time Interval", ylab="Average steps"  ,main="Daily Activity Pattern")
```

### Interval with max no. of steps.

```{r}
maxinterval<-subset(activity3,activity3$steps==max(activity3$steps),select=c(interval))
```
Interval with max no. of steps is `r maxinterval`.

## Imputing missing values.

### Total number of missing values in the dataset.

```{r}
NArows<-sum(is.na(activity))
```

No.of rows containing missing values in the dataset are `r NArows`.

### Imputing NA by mean of 5min interval and creting new dataset with imputed values.

```{r}
old<-read.csv("E:/Coursera/RWorkDir/activity.csv")
old2<-na.omit(activity)
library(plyr)
new<-ddply(old2,.(interval),summarise,
                 avg.steps=mean(as.numeric(steps)))
for(i in 1:17568)
{ 
  a<-NA
  if(old$steps[i] %in% a)
  {
    for(j in 1:288)
    {
      if(new$interval[j]==old$interval[i])
      {
        old$steps[i]<-new$avg.steps[j]
      }
    }
   }
}
```

A new dataset old is created.

```{r}
library(plyr)
Newactivity<-ddply(old,.(date),summarise,steps=sum(as.numeric(steps)))
library(lubridate)
Newactivity$date<-strptime(Newactivity$date,"%Y-%m-%d")
```

### Plot with new dataset i.e old after imputing NA.

```{r fig.width=11, fig.height=6}
plot(Newactivity$date,Newactivity$steps,type="h",lwd=11,col="dark red",lty=1,
     lend=1,main="No.of Steps taken per day(imputing NA by mean)",axes=F,xlab=NA,ylab=NA)
axis.POSIXct(1,format="%d %b",at=Newactivity$date,labels=TRUE,las=2,tck=-.01)
axis(side=2,las=2,tck=-.01)
mtext(side = 1, "Days", line = 4)
mtext(side = 2, "Steps", line = 3.2)
```

### Mean and Median
```{r}
Newstepsmean<-mean(old$steps)
```

The newmean of steps taken per day is `r Newstepsmean`(37.3826).

```{r}
Newstepsmedian<-median(old$steps)
```

The newmedian of steps taken per day is `r Newstepsmedian`(0).

After imputing the NA values by avg steps taken in 5min interval their is a significant difference in mean & median of both the datasets.

## Difference in between weekdays and weekends activity pattern.

```{r}
Newdata<-old
library(lubridate)
Newdata$date<-strptime(Newdata$date,"%Y-%m-%d")
Newdata$wd<-weekdays(Newdata$date)
for(i in 1:17568)
{
  if(Newdata$wd[i]=="Saturday"|Newdata$wd[i]=="Sunday")
  {
    Newdata$wd[i]<-"weekend"
  }
  else{
    Newdata$wd[i]<-"weekday"
  }
}
Newdata$wd<-factor(Newdata$wd)
```

```{r}
library(plyr)
Newdata1<-ddply(Newdata,.(interval,wd),summarise,steps=mean(as.numeric(steps)))
```

### Plot for weekdays & weekends.

```{r}
library(lattice)
xyplot(steps~interval|wd,data=Newdata1,layout=c(1,2),type="l")
```

